data wrangling and plots

### 2021-04-02 shared analysis files /data_wrangling_plot.Rmd
fixing up final plots to potentially include in results and discussion

---
title: "data wrangling and plots"
author: "Maggie Slein"
date: "3/8/21"
output: pdf_document
---

@@

#load data
dat_MA_0<- read_csv("metafor.csv") %>%
  mutate(resp_type = case_when( resp_units %in% c("1 day ^-1", "kJ *day^-1 *kg^-1", "m * s^-1") ~ "rate", TRUE ~ "trait"))

glimpse(dat_MA_0)

count(dat_MA_0, study_id, genus, species, mean_temp_constant, min_temp, max_temp)
```

@@

# Note in the above, you loose ~15% of data because of missing sample sizes and SEs reported, below these are removed for now
dat_MA_ES<-dat_MA_ES %>% 
  filter (!is.na(yi))

#removing outlier point to get better picture of data structure AND excluding CT assays 
normalized <- dat_MA_ES %>%
  filter(yi > -20) 
#filtering for just flux ranges of 10 C 
common_range <- dat_MA_ES %>% 
  filter(flux_range == "10")
#filtering for just studies with mean temp of 24
common_temp <- dat_MA_ES %>% 
  filter(mean_temp_constant == "24")
```

@@

## Relevant plots
### Figure 1.
```{r}
# boxplots of how fluctuation range influences SMD
ggplot(normalized, aes(x=flux_range, y=yi))+
  geom_point(alpha = 0.5)+
  theme_bw()+
  geom_smooth(method="lm", formula = y~x)+
  ggtitle("SMD across flux_range")
```

### Figure 2.
```{r}
# scatterplot of standardized mean response vs flux range colored and lm fit by org level
ggplot(normalized, aes(x=flux_range, y=yi, color = as.factor(org_level)))+
  geom_point()+
  geom_smooth(method="lm", formula = y~x)+
  theme_bw()+
  ggtitle("SMD across fluctuation ranges colored by organization level 
          and fit with linear model")
```

### Figure 3. 
```{r}
#boxplot of SMD across levels of organization
ggplot(normalized, aes(x=as.factor(org_level), y=yi, fill=as.factor(org_level)))+
  geom_boxplot(alpha =0.7)+
  geom_point(alpha = 0.3)+
  scale_fill_tron()+
  theme_bw()+
  theme(legend.position = "bottom")+
  ggtitle("SMD across organization level")
```

### Figure 4.
```{r}
#scatterplot of standardized mean difference across flux range colored by study fit with lm 
ggplot(normalized, aes(x=flux_range, y=yi, color = study_id))+
  geom_point()+
  geom_smooth(method="lm", formula = y~x)+
  theme_bw()+
  ggtitle("SMD across fluctuation ranges colored by 
          studies and fit with linear model by study")
```

@@

```{r, message=FALSE}
#corresponding random effects model 
fig1 <- rma.mv(yi, vi, data=dat_MA_ES, 
               random = ~1 | experiment_id/ study_id,
                 method="REML") 
fig1
#trying out mixed effects model 
fig1me <- rma(yi, vi, data=dat_MA_ES, mods = ~study_id,
                 method="FE")
fig1me
```

## Relevant Plots 
### Figure 5.
```{r}
#SMD across all studies
ggplot(normalized, aes(x=reorder(study_id, -yi), y=yi, color = as.factor(experiment_id)))+
  geom_boxplot()+
  scale_color_tron()+
  theme_minimal()+
  theme(axis.text.x = element_text(face = "bold", 
                                   size = 10, angle = 45,hjust = 1), 
        legend.position = "bottom")+
  ggtitle("SMD across all studies")
```

### Figure 6. 
```{r, echo= FALSE}
# scatterplot of standardized mean response faceted by exp age
ggplot(normalized, aes(x=yi, y=vi, color = study_id))+
  geom_point(alpha = 0.5)+
  theme_bw()+
  ggtitle("SMD across variance 
          colored by study")
```

### Figure 7.
```{r}
#looking at yi across studies with the same temperature range (10 C)
ggplot(common_range, aes(y=yi, x=reorder(study_id, -yi), color = study_id))+
  geom_boxplot()+
  geom_point(alpha = 0.5)+
  theme_minimal()+
  theme(axis.text.x = element_text(face = "bold", 
                                  size = 7, angle = 45))+
                                  ggtitle("SMD across studies with the same 
                                          temperature fluctuation range (10 C)")
```

### Figure 8.
```{r}
# scatterplot of how mean temperature influences SMD
ggplot(normalized, aes(x=mean_temp_constant, y=yi, color = study_id))+
  geom_point()+
  theme_minimal()+
  geom_smooth(formula = y~x, method = "lm")+
  ggtitle("SMD across mean temperature")
```

```{r}
#contour plot with flux_range on the x axis, mean temp of the y axis and colored by yi
fig <- plot_ly(
  x = normalized$flux_range, 
  y = normalized$mean_temp_constant, 
  z = normalized$yi, 
  type = "contour")
fig

@@

funnel_model <- rma.mv(yi, vi, data=normalized, mods = ~flux_range * mean_temp_constant +
                        exp_age + size + org_level, 
               random = ~1 | experiment_id/ study_id/ response_id,
                 method="REML") 
### draw funnel plots
#funnel plot with confidence intervals for significant values from different studies; taken from normalized data set because of big outlier 
funnel(funnel_model, level=c(90, 95, 99), shade=c("white", "gray55", "gray75"), refline=0, legend=TRUE)

?funnel

```

@@

@@

```{r, echo= FALSE, results = 'hide', message=FALSE, fig.show='hide'}
#trying to look at drosophila response across studies 
ggplot(drosophila, aes(y=yi, x=as.factor(flux_range), color = study_id))+
  geom_boxplot()+
  geom_point()+
  scale_color_tron()+
  theme_bw()+
  ggtitle("SMD across studies with the same study genus (Drosophila)")
#random effects model including org_level as a random variable
fig10 <- rma.mv(yi, vi, data=drosophila, mods = ~flux_range + study_id,
               random = ~1 | experiment_id/ study_id,
                 method="REML") 
fig10
#simple linear model looking at how flux_range and study_id affect yi in drosphila specific studies
simple10<-lm(yi~flux_range + study_id, data =drosophila)
summary(simple10)
```

@@

```{r, echo= FALSE, results = 'hide', message=FALSE, fig.show='hide'}
#plot across studies with common response variable body mass
ggplot(common_unit_g, aes(y=yi, color=as.factor(flux_range), x = study_id))+
  geom_boxplot()+
  theme_minimal()+
  theme(axis.text.x = element_text(face = "bold", 
                                  size = 7, angle = 45))+
  ggtitle("SMD across fluctuation ranges colored by studies with the common unit grams")
#random effects model within subset data
fig14 <- rma.mv(yi, vi, data=common_unit_g,
               random = ~1 | experiment_id/ study_id,
                 method="REML") 
fig14
#simple linear model looking at how flux_range affects yi within subset data
simple14<-lm(yi~study_id, data =common_unit_g)
summary(simple14)
```

>>>>

---
title: "data wrangling and plots"
author: "Maggie Slein"
date: "4/2/21"
output: pdf_document

---

@@

#load data
dat_MA_0<- read_csv("metafor.csv") %>%
  mutate(resp_type = case_when( resp_units %in% c("1 day ^-1", "kJ *day^-1 *kg^-1", "m * s^-1") ~ "rate", TRUE ~ "trait"))
```

@@

# Note in the above, you loose ~15% of data because of missing sample sizes and SEs reported, below these are removed for now
dat_MA_ES<-dat_MA_ES %>% 
  filter (!is.na(yi)) 
count(dat_MA_ES, study_id)

#removing outlier point to get better picture of data structure AND excluding CT assays 
normalized <- dat_MA_ES %>%
  filter(yi > -20) 
#filtering for just flux ranges of 10 C 
common_range <- dat_MA_ES %>% 
  filter(flux_range %in% c(10))
#filtering for just studies with mean temp of 24
common_temp <- dat_MA_ES %>% 
  filter(mean_temp_constant == "24")
```

@@

## Relevant plots

```{r, echo= FALSE}
# boxplots of how fluctuation range influences SMD
ggplot(normalized, aes(x=flux_range, y=yi))+
  geom_point(alpha = 0.5)+
  theme_bw()+
  geom_smooth(method="lm", formula = y~x)+
  ggtitle("SMD across flux_range")
```
**Figure 1.**
Data visualization supports model findings that there is a differene between shorter fluctuation ranges and longer fluctuation ranges.

```{r, echo= FALSE}
# scatterplot of standardized mean response vs flux range colored and lm fit by org level
ggplot(normalized, aes(x=flux_range, y=yi, color = as.factor(org_level)))+
  geom_point()+
  geom_smooth(method="lm", formula = y~x)+
  theme_bw()+
  ggtitle("SMD across fluctuation ranges colored by organization level 
          and fit with linear model")
```
**Figure 2.**
We can also see that there a difference in effect size in organization level responses across fluctuation range. Population level reponses trended positively across fluctuation ranges, while organism level responses trended negatively across fluctuation ranges.

```{r, echo= FALSE}
#scatterplot of standardized mean difference across flux range colored by study fit with lm 
ggplot(normalized, aes(x=flux_range, y=yi, color = study_id))+
  geom_point()+
  geom_smooth(method="lm", formula = y~x)+
  theme_bw()+
  ggtitle("SMD across fluctuation ranges colored by 
          studies and fit with linear model by study")
```
**Figure 3.**
When broken out by study, we can see that there is disagreement in the effect of fluctuation range on effect size.

```{r, echo= FALSE}
# scatterplot of standardized mean response vs flux range colored and lm fit by life history thermal characteristic
normalized %>%
  drop_na(stressful) %>%
ggplot(aes(x=flux_range, y=yi, color = stressful))+
  geom_point()+
  geom_smooth(method="lm", formula = y~x)+
  theme_bw()+
  ggtitle("SMD across fluctuation ranges colored by 
          whether or mean temperature was thermally stressful")
```
**Figure 4.** 
When we consider where the temperatures in the fluctuation range reach thermally stressful levels, there is a small difference between responses to stressful  temperatures (negative) and non-stressful temperatures (positive).

```{r, echo= FALSE}
# scatterplot of standardized mean response vs flux range colored and lm fit by life history thermal characteristic
normalized %>%
ggplot(aes(x=flux_range, y=yi, color = resp_type))+
  geom_point()+
  geom_smooth(method="lm", formula = y~x)+
  theme_bw()+
  ggtitle("SMD across fluctuation ranges colored by response type")
```
**Figure 5.**

@@

```{r, echo= FALSE, message=FALSE}
#corresponding random effects model 
fig1 <- rma.mv(yi, vi, data=dat_MA_ES, 
               random = ~1 | experiment_id/ study_id,
                 method="REML") 
fig1
```

## Relevant Plots

```{r, echo = FALSE}
#SMD across all studies
ggplot(normalized, aes(x=reorder(study_id, -yi), y=yi, color = as.factor(experiment_id)))+
  geom_boxplot()+
  scale_color_tron()+
  theme_minimal()+
  theme(axis.text.x = element_text(face = "bold", 
                                   size = 10, angle = 45,hjust = 1), 
        legend.position = "bottom")+
  ggtitle("SMD across all studies")
```

**Figure 6.**
When we look at experiment and study, we can see across the 15 studies included in this meta-analysis there are differences across and within experiments.

```{r, echo = FALSE}
#forest plot that shows the same thing?
forest <- normalized %>%
  group_by(study_id) %>%
  summarise(mean_yi = mean(yi),
            mean_vi = mean(vi)) %>%
  arrange(desc(mean_yi))
forest(forest$mean_yi, forest$mean_vi, slab=forest$study_id)
```

**Figure 7.**
I kind of like this figure better than the boxplot in Figure 5 as an overall figure?

```{r, echo = FALSE}
#looking at yi across studies with the same temperature range (10 C)
#excluded wang because there were too few points to do a boxplot
common_range %>%
  filter(study_id != "wang2015") %>%
ggplot(aes(y=yi, x=reorder(study_id, -yi), color = study_id))+
  geom_boxplot()+
  geom_point(alpha = 0.5)+
  theme_minimal()+
  theme(axis.text.x = element_text(face = "bold", 
                                  size = 7, angle = 45))+
                                  ggtitle("SMD across studies with the same 
                                          temperature fluctuation range (10 C)")
```

**Figure 8.**

```{r, echo = FALSE}
# scatterplot of how mean temperature influences SMD
ggplot(normalized, aes(x=mean_temp_constant, y=yi, color = study_id))+
  geom_point()+
  theme_minimal()+
  geom_smooth(formula = y~x, method = "lm")+
  ggtitle("SMD across mean temperature")
```

```{r, echo = FALSE}

@@

funnel_model <- rma.mv(yi, vi, data=normalized, mods = ~flux_range * mean_temp_constant +
                        exp_age + size + org_level, 
               random = ~1 | experiment_id/ study_id/ response_id,
                 method="REML") 
```

**Figure 9.**

```{r, echo = FALSE, fig.height=8, fig.width=12}
### draw funnel plots
#funnel plot with confidence intervals for significant values from different studies; taken from normalized data set because of big outlier 
funnel(funnel_model, level=c(90, 95, 99), shade=c("white", "gray55", "gray75"), refline=0, legend=TRUE)
funnel(funnel_model, yaxis="vi", main="Sampling Variance")
```

**Figure 10.** 
Well, from what I can gather, a majority of the effect sizes from our meta-analysis have a non-signficant effect size value. However, there is a population of a couple influential effect sizes that are significantly important.

@@

```{r, echo= FALSE, results = 'hide', message=FALSE, fig.show='hide'}
#trying out mixed effects model 
fig1me <- rma(yi, vi, data=dat_MA_ES, mods = ~study_id,
                 method="FE")
fig1me
```

```{r, echo= FALSE, results = 'hide', message=FALSE, fig.show='hide'}
#boxplot of SMD across levels of organization
ggplot(normalized, aes(x=as.factor(org_level), y=yi, fill=as.factor(org_level)))+
  geom_boxplot(alpha =0.7)+
  geom_point(alpha = 0.3)+
  scale_fill_tron()+
  theme_bw()+
  theme(legend.position = "bottom")+
  ggtitle("SMD across organization level")
```

```{r, echo= FALSE, results = 'hide', message=FALSE, fig.show='hide'}
# scatterplot of standardized mean response faceted by exp age
ggplot(normalized, aes(x=yi, y=vi, color = study_id))+
  geom_point(alpha = 0.5)+
  theme_bw()+
  ggtitle("SMD across variance 
          colored by study")
```

@@

```{r, echo = FALSE, eval= FALSE, results = 'hide', message=FALSE, fig.show='hide'}
#trying to look at drosophila response across studies 
ggplot(drosophila, aes(y=yi, x=as.factor(flux_range), color = study_id))+
  geom_boxplot()+
  geom_point()+
  scale_color_tron()+
  theme_bw()+
  ggtitle("SMD across studies with the same study genus (Drosophila)")
#random effects model including org_level as a random variable
fig10 <- rma.mv(yi, vi, data=drosophila, mods = ~flux_range + study_id,
               random = ~1 | experiment_id/ study_id,
                 method="REML") 
fig10
#simple linear model looking at how flux_range and study_id affect yi in drosphila specific studies
simple10<-lm(yi~flux_range + study_id, data =drosophila)
summary(simple10)
```

@@

```{r, echo = FALSE, eval= FALSE, results = 'hide', message=FALSE, fig.show='hide'}
#plot across studies with common response variable body mass
ggplot(common_unit_g, aes(y=yi, color=as.factor(flux_range), x = study_id))+
  geom_boxplot()+
  theme_minimal()+
  theme(axis.text.x = element_text(face = "bold", 
                                  size = 7, angle = 45))+
  ggtitle("SMD across fluctuation ranges colored by studies with the common unit grams")
#random effects model within subset data
fig14 <- rma.mv(yi, vi, data=common_unit_g,
               random = ~1 | experiment_id/ study_id,
                 method="REML") 
fig14
#simple linear model looking at how flux_range affects yi within subset data
simple14<-lm(yi~study_id, data =common_unit_g)
summary(simple14)
```
